Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children

### 2023-10-18 manuscript/main.tex
start on figure and table legend titles

\begin{sidewaysfigure}[!h]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
    \caption{
        \label{fig:1} Study design and overview. (\textbf{A}) Stool samples,
        cognitive assessments and neuroimaging were collected from participants
        at different ages throughout the first years of life.
        (\textbf{B}) Cognitive function scores are assessed by
        different instruments, but may be normalized using full-scale composite scores.
        (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on
        taxonomic profiles demonstrates high beta diversity, with much of the
        first axis of variation explained by increasing age and alpha diversity.
        Differences in gene function profiles (\textbf{E}) and neuroimaging (PCA based on
        the Euclidean distance of brain region volumes) (\textbf{F}) are likewise
        dominated by changes as children age. (\textbf{G}) Permutation analysis of
        variance (PERMANOVA) of taxonomic profiles, functional profiles
        (annotated as UniRef90s) and neuroimaging, against the metadata of
        interest. Variations in taxonomic and functional profiles explain a
        modest but significant percent of the variation in cognitive development
        in children over 18 months of age. (\textbf{H}) Mantel testing of different
        microbial feature matrices, shows overlapping but distinct patterns of
        variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
        used as cut-offs in some following models.
    }

@@

\begin{sidewaysfigure}[!htb]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure2.png}
    \caption{
        Taxa and gene functional groups are associated with cognitive function.
        (\textbf{A}) Volcano plot of multivariable linear model results showing the
        relationship between individual taxa and cognitive function score in
        children over 18 months of age (left) or across all ages (right),
        controlling for age, maternal education, and sequencing depth.
        Taxa that were significant after FDR correction (q
        \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
        associated with cognitive function,
        heatmaps of correlation with cognitive function in different age groups.
        (\textbf{C}) Multivariate linear models showing the relationship
        between individual taxa and domain-specific skill assessments
        from the MSEL for children under 6m (left), 18m to 36m (right),
        or over across all ages (right)
        controlling for age, maternal education, and sequencing depth.
        Asterisks in (A) and (C) indicate significance (q \textless{} 0.2)
        after multiple hypothesis correction.
        (\textbf{D-E}) Enrichment plots for proprionate degradation I gene set
        used in feature set enrichment analysis (FSEA) in children under 6m (D)
        or over 18m (E). (\textbf{G}) summary of FSEA
        results in children under 6m of age (left)
        over 18 months (middle) or across all ages (right), colored based on the
        significance of the association. Markers indicate the individual
        correlation of genes within a gene set,
        and vertical bars colored bars represent the
        median correlation of that gene set.
        Vertical dotted lines within each panel represent the median T statistic
        for all genes in that age group.
    }
    \label{fig:2}
\end{sidewaysfigure}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r|r|r|}
        \hline
        \textbf{Age group}& \multicolumn{2}{c|}{\textbf{0 to 6m}} & \multicolumn{2}{c|}{\textbf{18 to 120m}} \\\hline
        geneset & E.S. & q value & E.S. & q value \\\hline
        Acetate syn. I & -0.165 & 0.284 & 0.192 & \color{red}{\textbf{0.145}} \\
        Acetate syn. II & -0.491 & \color{blue}{\textbf{0.059}} & 0.248 & \color{red}{\textbf{0.15}} \\
        Acetate syn. III & -0.409 & \color{blue}{\textbf{0.133}} & 0.215 & 0.209 \\
        Butyrate syn. II & -0.357 & 0.284 & 0.414 & \color{red}{\textbf{0.178}} \\
        Glutamate syn. II & 0.343 & 0.284 & 0.355 & \color{red}{\textbf{0.023}} \\
        Propionate deg. I & -0.659 & \color{blue}{\textbf{0.029}} & 0.154 & \color{red}{\textbf{0.178}} \\\hline
    \end{tabular}
    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets.
    Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
    pseudomedian, respectively.}
    \end{center}
\end{table}

@@

\begin{sidewaysfigure}[!htb]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
    \caption{
        Random forest models predict both concurrent and future measured cognitive function.
        Comparison of RF predictor importance versus linear models for children
        over 18 months (\textbf{A}).
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model, if that species was significant (q
        \textless{} 0.2) in linear models, both, or neither.
        Comparison of RF predictor importance for microbial features
        between concurrent and future cognitive function measure
        (\textbf{B}).
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model only for concurrent prediction,
        only for the future prediction, or for both models.
        Features absent of one model are shown on the origin-aligned sideplots.
        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
        Blue violin plot shows distribution of sampling age,
        while red violin plot shows distribution of asessment age.
        Black lines connect ages for each individual subject.      
        Ranked feature importance for taxa in RF models for
        both concurrent (\textbf{D}) and future (\textbf{E})
        cognitive function measurement prediction.
        Red dots indicate how relative importances change (among taxa)
        when demographics (sex, SES) are arred as covariables.
        Abundances and prevalences of taxa that are important
        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
        }
    \label{fig:3}
\end{sidewaysfigure}

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r|}
      \hline\hline
      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
      \multirow{5}{*}{0 to 6} & -    & + & 13.37 ± 0.05 & -0.16 ± 0.01 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 12.99 ± 0.04 & -0.11 ± 0.01 \\ \cline{3-5}
            &                        & + & 13.01 ± 0.04 & -0.13 ± 0.01 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{3-5}
            &                        & + & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{1-5}
      \multirow{5}{*}{18 to 120} & - & + & 17.06 ± 0.04 & 0.511 ± 0.003 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 18.71 ± 0.04 & 0.347 ± 0.003 \\ \cline{3-5}
            &                        & + & 18.25 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 18.96 ± 0.04 & 0.299 ± 0.003 \\ \cline{3-5}
            &                        & + & 18.74 ± 0.04 & 0.341 ± 0.003 \\ \hline\hline
    \end{tabular}
    \caption{\label{tab:rfbench}Benchmark metrics for the cognitive assessment score
    prediction models. Confidence intervals (C.I.) are calculated from the
    distribution of metrics from repeated CV at a confidence level of 95.
    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
    \end{center}
\end{table}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r}
        \hline
        \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set Root-mean-square-error (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
        - & + & 20.25 ± 0.07 & 0.43 ± 0.0 \\ \cline{1-4}
        \multirow{2}{*}{taxa} & - & 22.32 ± 0.08 & 0.07 ± 0.0 \\ \cline{2-4}
            & + & 21.53 ± 0.08 & 0.27 ± 0.0 \\ \cline{1-4}
        \multirow{2}{*}{genes} & - & 22.21 ± 0.08 & 0.12 ± 0.0 \\ \cline{2-4}
            & + & 22.08 ± 0.08 & 0.15 ± 0.0 \\\hline\hline
    \end{tabular}
    \caption{\label{tab:rfbench2}Benchmark metrics for the cognitive assessment score
    future prediction models. Confidence intervals (C.I.) are calculated from the
    distribution of metrics from repeated CV at a confidence level of 95.
    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
    \end{center}
\end{table}

@@

\begin{table}[!h]
    \begin{center}
        \begin{tabular}{|r|r|r|r|r|}
        \hline
        \textbf{Target Mullen scale} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
        \multirow{5}{*}{Expressive Language} & - & + & 11.17 ± 0.04 & 0.48 ± 0.0 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 12.37 ± 0.04 & 0.25 ± 0.0 \\ \cline{3-5}
            &                        & + & 12.05 ± 0.04 & 0.37 ± 0.0 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 12.59 ± 0.04 & 0.15 ± 0.0 \\ \cline{3-5}
            &                        & + & 12.46 ± 0.04 & 0.21 ± 0.0 \\ \cline{1-5}
        \multirow{5}{*}{Gross Motor} & - & + & 10.53 ± 0.04 & 0.346 ± 0.006 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 11.08 ± 0.04 & 0.134 ± 0.006 \\ \cline{3-5}
            &                        & + & 11.07 ± 0.04 & 0.137 ± 0.006 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \cline{3-5}
            &                        & + & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \hline
        \multirow{5}{*}{Visual reception} & - & + & 14.13 ± 0.04 & 0.462 ± 0.003 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 15.49 ± 0.04 & 0.245 ± 0.004 \\ \cline{3-5}
            &                        & + & 15.2 ± 0.04 & 0.333 ± 0.004 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 15.97 ± 0.04 & 0.067 ± 0.004 \\ \cline{3-5}
            &                        & + & 15.84 ± 0.04 & 0.118 ± 0.005 \\ \hline\hline
        \end{tabular}
        \caption{\label{tab:subscalebenchmark}Benchmark metrics for the Mullen subscales prediction models.
        Confidence intervals (C.I.) are calculated from the
        distribution of metrics from repeated CV at a confidence level of 95.
        Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
        Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.
        Age range, in all cases, was 18-120 months.}
    \end{center}
\end{table}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|}
      \hline\hline
      \textbf{Statistic} & \textbf{MAPE} & \textbf{R} \\\hline
      Mean & 0.073 & 0.146 \\
      Standard deviation & 0.029 & 0.131 \\
      Maximum & 0.205 & 0.598 \\
      75th percentile & 0.081 & 0.212 \\
      50th percentile & 0.066 & 0.129 \\
      25th percentile & 0.057 & 0.073 \\
      Minimum & 0.036 & -0.133 \\\hline
    \end{tabular}
    \caption{\label{tab:neuro}Summary statistics for the neuroanatomy prediction benchmarks.
    Mean absolute proportional error (MAPE) and correlation coeficient (R)
    of test-set preditions are reported accross models for all brain regions.
    Complete list of values can be found on Table S6.}
    \end{center}
\end{table}

@@

\begin{sidewaysfigure}[!htb]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure4.png}
    \caption{
        Microbial feature importance in predicting MSEL subscores and brain volumes in children over 18 months of age.
        (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,
        with or without demographics included as predictors.
        (\textbf{B}) Important taxa in models predicting MSEL subscale scores.
        (\textbf{C}) Average test-set correlations for prediction of MRI segmentation
        data from microbiome and demographics on select segments after repeated
        cross validation. (\textbf{D}) Heatmap of average individual relative taxa
        importances on each brain segment. Importances are reported as
        proportions relative to the sum of importances for each model - since
        every model is trained with 132 features, and even distribution of
        importance would be 0.75\% for each feature. Segments and taxa ordered
        by HCA on a select list of species with high ranks on average importances,
        and their respective highest-load segments (\textbf{E}).
    }
    \label{fig:4}
\end{sidewaysfigure}

>>>>

\begin{sidewaysfigure}[!h]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
    \caption{
        \label{fig:1} \textbf{The RESONANCE cohort is a diverse cohort with longitudinal sampling.} (\textbf{A}) Stool samples,
        cognitive assessments and neuroimaging were collected from participants
        at different ages throughout the first years of life.
        (\textbf{B}) Cognitive function scores are assessed by
        different instruments, but may be normalized using full-scale composite scores.
        (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on
        taxonomic profiles demonstrates high beta diversity, with much of the
        first axis of variation explained by increasing age and alpha diversity.
        Differences in gene function profiles (\textbf{E}) and neuroimaging (PCA based on
        the Euclidean distance of brain region volumes) (\textbf{F}) are likewise
        dominated by changes as children age. (\textbf{G}) Permutation analysis of
        variance (PERMANOVA) of taxonomic profiles, functional profiles
        (annotated as UniRef90s) and neuroimaging, against the metadata of
        interest. Variations in taxonomic and functional profiles explain a
        modest but significant percent of the variation in cognitive development
        in children over 18 months of age. (\textbf{H}) Mantel testing of different
        microbial feature matrices, shows overlapping but distinct patterns of
        variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
        used as cut-offs in some following models.
    }

@@

\begin{sidewaysfigure}[!htb]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure2.png}
    \caption{
        \textbf{Taxa and gene functional groups are associated with cognitive function.}
        (\textbf{A}) Volcano plot of multivariable linear model results showing the
        relationship between individual taxa and cognitive function score in
        children over 18 months of age (left) or across all ages (right),
        controlling for age, maternal education, and sequencing depth.
        Taxa that were significant after FDR correction (q
        \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
        associated with cognitive function,
        heatmaps of correlation with cognitive function in different age groups.
        (\textbf{C}) Multivariate linear models showing the relationship
        between individual taxa and domain-specific skill assessments
        from the MSEL for children under 6m (left), 18m to 36m (right),
        or over across all ages (right)
        controlling for age, maternal education, and sequencing depth.
        Asterisks in (A) and (C) indicate significance (q \textless{} 0.2)
        after multiple hypothesis correction.
        (\textbf{D-E}) Enrichment plots for proprionate degradation I gene set
        used in feature set enrichment analysis (FSEA) in children under 6m (D)
        or over 18m (E). (\textbf{G}) summary of FSEA
        results in children under 6m of age (left)
        over 18 months (middle) or across all ages (right), colored based on the
        significance of the association. Markers indicate the individual
        correlation of genes within a gene set,
        and vertical bars colored bars represent the
        median correlation of that gene set.
        Vertical dotted lines within each panel represent the median T statistic
        for all genes in that age group.
    }
    \label{fig:2}
\end{sidewaysfigure}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r|r|r|}
        \hline
        \textbf{Age group}& \multicolumn{2}{c|}{\textbf{0 to 6m}} & \multicolumn{2}{c|}{\textbf{18 to 120m}} \\\hline
        geneset & E.S. & q value & E.S. & q value \\\hline
        Acetate syn. I & -0.165 & 0.284 & 0.192 & \color{red}{\textbf{0.145}} \\
        Acetate syn. II & -0.491 & \color{blue}{\textbf{0.059}} & 0.248 & \color{red}{\textbf{0.15}} \\
        Acetate syn. III & -0.409 & \color{blue}{\textbf{0.133}} & 0.215 & 0.209 \\
        Butyrate syn. II & -0.357 & 0.284 & 0.414 & \color{red}{\textbf{0.178}} \\
        Glutamate syn. II & 0.343 & 0.284 & 0.355 & \color{red}{\textbf{0.023}} \\
        Propionate deg. I & -0.659 & \color{blue}{\textbf{0.029}} & 0.154 & \color{red}{\textbf{0.178}} \\\hline
    \end{tabular}
    \caption{\label{tab:fsea} \textbf{Neuroactive microbial gene sets microbial gene sets are associated with cognitive function}.
    Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
    pseudomedian, respectively.}
    \end{center}
\end{table}

@@

\begin{sidewaysfigure}[!htb]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
    \caption{
        \textbf{Random forest models predict both concurrent and future measured cognitive function.}
        Comparison of RF predictor importance versus linear models for children
        over 18 months (\textbf{A}).
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model, if that species was significant (q
        \textless{} 0.2) in linear models, both, or neither.
        Comparison of RF predictor importance for microbial features
        between concurrent and future cognitive function measure
        (\textbf{B}).
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model only for concurrent prediction,
        only for the future prediction, or for both models.
        Features absent of one model are shown on the origin-aligned sideplots.
        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
        Blue violin plot shows distribution of sampling age,
        while red violin plot shows distribution of asessment age.
        Black lines connect ages for each individual subject.      
        Ranked feature importance for taxa in RF models for
        both concurrent (\textbf{D}) and future (\textbf{E})
        cognitive function measurement prediction.
        Red dots indicate how relative importances change (among taxa)
        when demographics (sex, SES) are arred as covariables.
        Abundances and prevalences of taxa that are important
        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
        }
    \label{fig:3}
\end{sidewaysfigure}

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r|}
      \hline\hline
      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
      \multirow{5}{*}{0 to 6} & -    & + & 13.37 ± 0.05 & -0.16 ± 0.01 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 12.99 ± 0.04 & -0.11 ± 0.01 \\ \cline{3-5}
            &                        & + & 13.01 ± 0.04 & -0.13 ± 0.01 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{3-5}
            &                        & + & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{1-5}
      \multirow{5}{*}{18 to 120} & - & + & 17.06 ± 0.04 & 0.511 ± 0.003 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 18.71 ± 0.04 & 0.347 ± 0.003 \\ \cline{3-5}
            &                        & + & 18.25 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 18.96 ± 0.04 & 0.299 ± 0.003 \\ \cline{3-5}
            &                        & + & 18.74 ± 0.04 & 0.341 ± 0.003 \\ \hline\hline
    \end{tabular}
    \caption{\label{tab:rfbench} \textbf{RF models on microbial taxa and genes can learn associations with concurrently measured cognitive performance.}
    Benchmark metrics for the cognitive assessment score
    prediction models. Confidence intervals (C.I.) are calculated from the
    distribution of metrics from repeated CV at a confidence level of 95.
    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
    \end{center}
\end{table}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|r|r}
        \hline
        \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set Root-mean-square-error (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
        - & + & 20.25 ± 0.07 & 0.43 ± 0.0 \\ \cline{1-4}
        \multirow{2}{*}{taxa} & - & 22.32 ± 0.08 & 0.07 ± 0.0 \\ \cline{2-4}
            & + & 21.53 ± 0.08 & 0.27 ± 0.0 \\ \cline{1-4}
        \multirow{2}{*}{genes} & - & 22.21 ± 0.08 & 0.12 ± 0.0 \\ \cline{2-4}
            & + & 22.08 ± 0.08 & 0.15 ± 0.0 \\\hline\hline
    \end{tabular}
    \caption{\label{tab:rfbench2}  \textbf{RF models on microbial taxa and genes can learn associations with future cognitive performance.}
    Benchmark metrics for the cognitive assessment score
    future prediction models. Confidence intervals (C.I.) are calculated from the
    distribution of metrics from repeated CV at a confidence level of 95.
    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
    \end{center}
\end{table}

@@

\begin{table}[!h]
    \begin{center}
        \begin{tabular}{|r|r|r|r|r|}
        \hline
        \textbf{Target Mullen scale} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
        \multirow{5}{*}{Expressive Language} & - & + & 11.17 ± 0.04 & 0.48 ± 0.0 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 12.37 ± 0.04 & 0.25 ± 0.0 \\ \cline{3-5}
            &                        & + & 12.05 ± 0.04 & 0.37 ± 0.0 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 12.59 ± 0.04 & 0.15 ± 0.0 \\ \cline{3-5}
            &                        & + & 12.46 ± 0.04 & 0.21 ± 0.0 \\ \cline{1-5}
        \multirow{5}{*}{Gross Motor} & - & + & 10.53 ± 0.04 & 0.346 ± 0.006 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 11.08 ± 0.04 & 0.134 ± 0.006 \\ \cline{3-5}
            &                        & + & 11.07 ± 0.04 & 0.137 ± 0.006 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \cline{3-5}
            &                        & + & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \hline
        \multirow{5}{*}{Visual reception} & - & + & 14.13 ± 0.04 & 0.462 ± 0.003 \\ \cline{2-5}
            & \multirow{2}{*}{taxa}  & - & 15.49 ± 0.04 & 0.245 ± 0.004 \\ \cline{3-5}
            &                        & + & 15.2 ± 0.04 & 0.333 ± 0.004 \\ \cline{2-5}
            & \multirow{2}{*}{genes} & - & 15.97 ± 0.04 & 0.067 ± 0.004 \\ \cline{3-5}
            &                        & + & 15.84 ± 0.04 & 0.118 ± 0.005 \\ \hline\hline
        \end{tabular}
        \caption{\label{tab:subscalebenchmark}  \textbf{RF models on microbial taxa and genes can learn associations with performance in cognitive subdomains.}
        Benchmark metrics for the Mullen subscales prediction models.
        Confidence intervals (C.I.) are calculated from the
        distribution of metrics from repeated CV at a confidence level of 95.
        Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
        Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.
        Age range, in all cases, was 18-120 months.}
    \end{center}
\end{table}

@@

\begin{table}[!h]
    \begin{center}
    \begin{tabular}{|r|r|r|}
      \hline\hline
      \textbf{Statistic} & \textbf{MAPE} & \textbf{R} \\\hline
      Mean & 0.073 & 0.146 \\
      Standard deviation & 0.029 & 0.131 \\
      Maximum & 0.205 & 0.598 \\
      75th percentile & 0.081 & 0.212 \\
      50th percentile & 0.066 & 0.129 \\
      25th percentile & 0.057 & 0.073 \\
      Minimum & 0.036 & -0.133 \\\hline
    \end{tabular}
    \caption{\label{tab:neuro}  \textbf{RF models on microbial taxa predicting brain regions have generally good performance.}
    Mean absolute proportional error (MAPE) and correlation coeficient (R)
    of test-set preditions are reported accross models for all brain regions.
    Complete list of values can be found on Table S6.}
    \end{center}
\end{table}

@@

\begin{sidewaysfigure}[!htb]
    \centering
    \small
    \includegraphics[width=0.9\textwidth]{assets/Figure4.png}
    \caption{
        \textbf{Microbial features can train RF models to predict MSEL subscores and brain volumes in children over 18 months of age}.
        (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,
        with or without demographics included as predictors.
        (\textbf{B}) Important taxa in models predicting MSEL subscale scores.
        (\textbf{C}) Average test-set correlations for prediction of MRI segmentation
        data from microbiome and demographics on select segments after repeated
        cross validation. (\textbf{D}) Heatmap of average individual relative taxa
        importances on each brain segment. Importances are reported as
        proportions relative to the sum of importances for each model - since
        every model is trained with 132 features, and even distribution of
        importance would be 0.75\% for each feature. Segments and taxa ordered
        by HCA on a select list of species with high ranks on average importances,
        and their respective highest-load segments (\textbf{E}).
    }
    \label{fig:4}
\end{sidewaysfigure}
